---
title: "project1"
author: "Thrinesh Duvvuru"
date: "24/10/2020"
output: html_document
---

REPRODUCIBLE RESEARCH PROJECT 1
===============================

##Unzipping the data
```{r}
unzip(zipfile="repdata_data_activity.zip",exdir = getwd())
```

##loading activity data
```{r}
activity<-read.csv("activity.csv",header = TRUE)
```

##Mean total number of steps taken per day
```{r}
steps<-tapply(activity$steps,activity$date,sum)
hist(steps,xlab="number of steps",main="steps per day")
```

##Average daily activity pattern
```{r}
averagesteps<-tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
plot(as.numeric(names(averagesteps)),averagesteps,xlab="interval",ylab="steps",main="Average daily activity",type = "l")
```

##The 5 min interval that contain maximum number of steps
```{r}
maxinterval<-names(sort(averagesteps,decreasing = TRUE)[1])
maxsteps<-sort(averagesteps,decreasing = TRUE)[1]
```

##Calculating the number of missing values
```{r}
missingvalues<-sum(is.na(activity$steps))
```

##Filling the missing values with mean and creating new dataset
```{r}
averagesteps<-tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
splitactivity<-split(activity,activity$interval)
for (i in 1:length(splitactivity)) {
  splitactivity[[i]]$steps[is.na(splitactivity[[i]]$steps)]<-averagesteps[i]
}
imputedactivity<-do.call("rbind",splitactivity)
imputedactivity<-imputedactivity[order(imputedactivity$date),]
```

##Total number of steps taken each day,mean and median
```{r}
Totalsteps<-tapply(imputedactivity$steps,imputedactivity$date,sum)
hist(Totalsteps,xlab="numberof steps",main="steps per day (imputed data)")
```  
```{r}
Mean.imputed<-mean(Totalsteps,na.rm = TRUE)
median.imputed<-median(Totalsteps,na.rm = TRUE)
```

##Are there differences in activity patterns between weekdays and weekends?
```{r}
imputedactivity$day<-ifelse(weekdays(as.Date(imputedactivity$date))=="Saturday"|weekdays(as.Date(imputedactivity$date))=="Sunday","weekend","weekday")
```

##Making a panel plot containing a time series plot of 5 minute interval and average number of steps taken,averaged across all weekends and weekdays
```{r}
Steps.weekend <-tapply(imputedactivity[imputedactivity$day=="weekend",]$steps,imputedactivity[imputedactivity$day=="weekend",]$interval, mean,na.rm=TRUE)

steps.weekday<-tapply(imputedactivity[imputedactivity$day=="weekday",]$steps,imputedactivity[imputedactivity$day=="weekday",]$interval, mean,na.rm=TRUE)
par(mfrow=c(1,2))
plot(as.numeric(names(Steps.weekend)),Steps.weekend,xlab="interval",ylab="steps",main="weekend activity pattern",type = "l")
plot(as.numeric(names(steps.weekday)),steps.weekday,xlab="interval",ylab="steps",main="weekdays activity pattern",type = "l")
```








